Supplementary Materials

### 2023-06-17 manuscript/manuscript.tex
ms changes

Our method takes tracks of annotated features (an ``annotation model") that are
\emph{a priori} expected to have similar fitness effects, and estimates the
distribution of fitness effects for each feature type. We consider two classes
of annotation models: (1) CADD score models, which assign a rank-order
pathogenicity score to each basepair, and (2) and more interpretable gene
feature-based models that includes protein coding regions, intron and UTRs, and
PhastCons regions. We include PhastCons regions because there is abundant
evidence of highly-conserved non-coding regions in humans that we would expect
to be under negative selection and impact patterns of linked selection (XXX).
Since our method fits a DFE for each feature type, PhastCons regions that
overlap other gene annotations (e.g. coding sequences) must be assigned to
either of the two feature types. Since this can impact estimates, we fit both
alternatives: a PhastCons Priority model, where features that overlap PhastCons
regions are classified as PhastCons, and a Feature Priority model, where gene
structure features take priority and the PhastCons type catches all
highly-conserved non-genic regions. In total, we fit four annotation models
(CADD 6\%, CADD 8\%, PhastCons Priority, and Feature Priority) to high-coverage
1000 Genome data for three populations: Yoruba (YRI), Han Chinese (CHB), and
European (CEU) ancestries.

@@

Since our $R_\text{LOO}^2$ estimates suggest our models fit left-out
chromosomes well even though our method assumes constant demography and
homogeneous mutation rates along the genome, we wondered what $R^2$ we could
expect under a simple theoretic model where $\cov_i(B_i', \varepsilon_i) = 0$
and $\var_i(\varepsilon_i)$ was determined by the variance in coalescence times
due to drift alone with an effective population size rescaled by $B_i'$. We
calculate this ``drift $R^2$" by calculating the theoretic coalescence time
variance in a megabase window with recombination and an effective population
size of $B_i' N_e$ where we use our best-fitting model's as a plug-in estimator
of $B_i'$. We note that this is a rough approximation, since this calculation
assumes constant demography and that the variance in coalescence times under
negative selection is likely lower than expected under drift alone.

Still, we find that our out-sample $R_\text{LOO}^2$ for Yoruba individuals
($R_\text{LOO}^2 \approx 68$\%) is close to the theoretic drift
$R_\text{drift}^2 \approx 67$\%. For bottlenecked out-of-Africa populations,
$R_\text{drift}^2 \approx 64$\% compared to the observed $R_\text{LOO}^2
\approx 62$\% for CEU and $R_\text{LOO}^2 \approx 59$\% for CHB. We note that
because bottlenecks increase the variance in coalescence times beyond the level
implied by the effective population size, the disagreement between observed
$R_\text{LOO}^2$ and the theoretic $R_\text{drift}^2$ is likely to shrink.
Overall, this suggests that negative selection models explain the vast majority
coalescence time variation at the megabase-scale that is capable of being
explained (i.e. that is not coalescence noise).

@@

@@

\subsection*{DFE Estimates}

Our best-fitting models have out-sample $R^2$ of around 68\% for Yoruba, 62\%
European, and 60\% for Han Chinese ancestry individuals. However, $R^2$ has a
theoretic upper bound due to coalescence noise around the predicted mean level
$\widehat{\pi}(x)$.

>>>>

Our method takes tracks of annotated features (an ``annotation model") that are
\emph{a priori} expected to have similar fitness effects, and estimates the
distribution of fitness effects for each feature type. We consider two classes
of annotation models: (1) CADD score models, which assign a rank-order
pathogenicity score to each basepair, and (2) and more interpretable gene
feature-based models that includes protein coding regions, intron and UTRs, and
PhastCons regions. We include PhastCons regions because there is abundant
evidence of highly-conserved non-coding regions in humans that we would expect
to be under negative selection and impact patterns of linked selection (XXX).
Since our method fits a DFE for each feature type, PhastCons regions that
overlap other gene annotations (e.g. coding sequences) must be assigned to
either of the two feature types. Since this can impact estimates, we fit both
alternatives: a PhastCons Priority model, where features that overlap PhastCons
regions are classified as PhastCons, and a Feature Priority model, where gene
structure features take priority and the PhastCons type catches all
highly-conserved non-genic regions. In total, we fit four annotation models
(CADD 6\%, CADD 8\%, PhastCons Priority, and Feature Priority) to high-coverage
1000 Genome data for three populations: Yoruba (YRI), Han Chinese (CHB), and
European (CEU) ancestries. We assessed and compared models according to how
well they predict the genome-wide patterns of diversity they were fit on. We
use $R_\text{LOO}^2$, which is the proportion of the observed variance in
megabase-scale diversity that is predicted by our model in out-samples (i.e.
whole chromosomes left out during the fitting process).

@@

Since our $R_\text{LOO}^2$ estimates suggest our models fit left-out
chromosomes well even though our method assumes constant demography and
homogeneous mutation rates along the genome, we wondered what $R^2$ we could
expect under a simple theoretic model where $\cov_i(B_i', \varepsilon_i) = 0$
and $\var_i(\varepsilon_i)$ was determined by the variance in neutral
coalescence times alone with an effective population size rescaled by $B_i'$.
We calculate this ``coalescence $R^2$" by calculating the theoretic coalescence
time variance in a megabase window with recombination and an effective
population size of $B_i' N_e$ where we use our best-fitting model's as a
plug-in estimator of $B_i'$. We note that this is a rough approximation, since
this calculation assumes constant demography and that the variance in
coalescence times under negative selection is likely lower than expected under
drift alone.

Still, we find that our out-sample $R_\text{LOO}^2$ for Yoruba individuals
($R_\text{LOO}^2 \approx 68$\%) is close to the theoretic $R_\text{coal}^2
\approx 67$\%. For bottlenecked out-of-Africa populations, $R_\text{coal}^2
\approx 64$\% compared to the observed $R_\text{LOO}^2 \approx 62$\% for CEU
and $R_\text{LOO}^2 \approx 59$\% for CHB. We note that because bottlenecks
increase the variance in coalescence times beyond the level implied by the
effective population size, the disagreement between observed $R_\text{LOO}^2$
and the theoretic $R_\text{coal}^2$ is likely to shrink. Overall, this suggests
that negative selection models explain the vast majority coalescence time
variation at the megabase-scale that is capable of being explained (i.e. that
is not coalescence noise).

@@

\begin{figure}[htbp] \centering
    \includegraphics{figures/figure_4_pop_mu_trimmed.pdf} 
    \caption{}
  \label{fig:figure-4}
\end{figure}

@@

\subsection*{Predicted Diversity and $B'(x)$ maps}

When we fit our negative selection model to genome-wide patterns of diversity,
the parameter estimates generate three predicted quantities: (1) the $B(x)$
map, (2) the substitution rates per segment, and (3) the fitness variance per
segment. Here we focus on the predicted $B(x)$ map and implied levels of
diversity at the megabase scale. We discuss substitution rates and fitness
variance in the next section.

We also obtain a prediction $\widehat{B}_i$ for the reduction factor $B_i$ in
each window $i$, which can be multiplied by our estimate of $\widehat{\pi}_0$
to obtain a prediction of diversity $\widehat{\pi}_i = \widehat{\pi}_0
\widehat{B}_i$.

\subsection*{Model Predictions and the Substitution Process}

Our method accommodates weak selection by modeling the fitness variance in each
segment under an equilibrium negative selection process. This fitness variance
is found by simultaneously solving for the per-generation substitution rate $R$
and the draft-effective population size $N_d$ for each segment. When our model
is fit to genome-wide diversity data, the parameter estimates imply predictions
of the deleterious substitution rate per basepair, per generation, for each
feature. Our method only considers within-species pairwise diversity, so these
predicted substitution rates constitute a test that our model is making
reasonable out-sample predictions of the substitution rates across features.

We focus specifically on the predictions from our feature-based annotation
models, since we can readily estimate human substitution distances for each
feature to compare to our predictions.

The estimated substitution distance should in theory equal the product of the
substitution rate per site, per year and the branch length in years.

When comparing our predicted mutation rates to observed substitution distances,
we must account for three major sources of uncertainty.

First, the split time between humans and chimpanzees is unknown, with estimates
ranging from 6Mya to 12Mya (XXX). Second, our predictions are in sites per
generation, and the appropriate human generation time is unknown. Third,

For a particular feature $f$, the estimated per-site substitution distance
$S_f$ along a branch (which accounts for back mutations) should correspond to
$S_f = T_g r_f$, where $T_g$ is the time in generations and $r_f$ is the
substitution rate per basepair, per generation of that feature.

Each feature is
composed of basepairs with varying selection coefficients distributed according
to the feature-specific DFE vector $\mathbf{w}_f$ and corresponding vector of
substitution rates per selection coefficient, $\mathbf{r}_f$.

Our estimate of a feature's DFE
is $\widehat{\mathbf{w}}_f$.

We calculate the

\begin{align}
    \label{eq:sub}
    r_f = \mu \int_0^1 p_f(s) ds
\end{align}

Solving the XXX equations, we end up with predictions $\widehat{r}_f$ for the
per-basepair, per-generation substitution rate for each feature. These are
calibrated to the MLE mutation rate estimate $\widehat{\mu}_\text{del}$, which
as mentioned, excludes the contribution of lethal mutations which act purely to
reduce $N_e$ and thus only impact $\pi_0$. Using Equation \eqref{eq:sub} we can
re-calibrate these estimates against other mutation rates.

Because of it's greater interpretabilty, we initially focus on the substitution
rate estimates from the Feature Priority model even though it had a poorer-fit
and high mutation rate estimate.

We can also re-calibrate this model to an alternate mutation rate.

### 2023-06-17 manuscript/supplementary.tex
ms changes

>>>>

\section*{Ratchet Rate Predictions}

\section*{Load Predictions}

Each fixation of a mutant with effect $-s$ decreases the population fitness by
$(1-2s)$ since we assume additive effects throughout. For each feature's DFE
vector $\mathbf{w}$ and grid of selection coefficients $\mathbf{s}$, we have a
corresponding vector $\widehat{\mathbf{r}}$ of predicted ratchet rates. Each
generation $\widehat{\mathbf{r}}$ substitutions are fixed each generation.
Since we assume multiplicative fitness, the log load for selection coefficient
$i$ in the vector is,

\begin{align}
    l_i = \log(1-2 s_i) \widehat{r_i}.
\end{align}

The total log load is,

\begin{align}
    \log(L) = \sum_{i} l_i.
\end{align}
